# Analysis of Data Carpentry's Archived Survey Data
__Contributors__: [Kari L. Jordan](https://github.com/kariljordan) and [Paula Andrea](https://github.com/orchid00)
__Work Cycle__: Ganymede

```{r setup, include=FALSE}
library(knitr)
opts_chunk$set(echo = FALSE,
               message = FALSE,
               warning = FALSE)
library(tidyverse)
library(ggplot2)
library(dplyr)
library(tidyr)
library(readr)
library(purrr)
library(tibble)
library(DBI)
library(ggmap)
library(likert)
library(mapproj)
library(RColorBrewer)
library(forcats)
```
In August 2017 Data Carpentry launched new pre- and post-workshop surveys that included skills-based questions. This report is an analysis of the old pre- and post-workshop surveys. 

Responses collected for Data Carpentry's workshops from XX through XX are included in this report. In this XX year period, nearly XX responses were collected.

PDFs of both the [pre](https://github.com/carpentries/assessment-projects/blob/master/data-carpentry-projects/presurvey.pdf) and [post](https://github.com/carpentries/assessment-projects/blob/master/data-carpentry-projects/postsurvey.pdf) workshop surveys, the data used in this analysis, and full R code are located in the [data-carpentry-projects](https://github.com/carpentries/assessment-projects/tree/master/data-carpentry-projects) folder in the Carpentries [assessment-projects](https://github.com/carpentries/assessment-projects) repo on GitHub. 

Community members are invited to contribute code to this analysis. Feel free to use the data and [tell us](mailto: kariljordan@carpentries.org) about your findings.

```{r include=FALSE}
# KLJ Load the pre- and post-workshop data into R
predata <- readr::read_csv("https://raw.githubusercontent.com/carpentries/assessment-projects/master/data-carpentry-projects/preworkshop-public-archived.csv")

postdata <- readr::read_csv("https://raw.githubusercontent.com/carpentries/assessment-projects/master/data-carpentry-projects/postworkshop-public-archived.csv")
```
### Pre-Workshop Survey Demographics

A breakdown of Data Carpentry's learners (pre-workshop) by __status__ is provided below.
```{r echo=FALSE}
### KLJ Plot for pre-survey respondents status
prestatus = c(
    "Undergraduate Student",
    "Graduate Student",
    "Post-doc",
    "Faculty",
    "Industry",
    "Staff",
    "Other (please specify)"
  )
prestatus = factor(prestatus)

predata$Status = factor(predata$Status, levels = prestatus)

data_prestatus_tally <- 
  predata %>% 
  group_by(Status) %>% 
  tally() %>% 
  filter(!is.na(Status)) 

ggplot(data_prestatus_tally, 
       aes(Status, y = 100 * (n/sum(n)),
           n)) +
  geom_bar(stat = "identity", fill="orange") +
  geom_text(aes(label=n), size= 4) +
  scale_x_discrete(labels = function(x) lapply(strwrap(x, width = 10, simplify = FALSE), paste, collapse="\n")) +
  theme_classic() +
  xlab("Status (Pre-Survey)") +
  ylab("% Respondents") +
  ggtitle("Majority of Pre-Survey Respondents were Graduate Students") +
  theme(plot.title = element_text(hjust = 0.5)) +
  theme_classic(base_size = 14)


### KLJ Plot for post-survey respondents status
poststatus = c(
    "Undergraduate Student",
    "Graduate Student",
    "Post-doc",
    "Faculty",
    "Industry",
    "Staff",
    "Other (please specify)"
  )
poststatus = factor(poststatus)

postdata$Status = factor(postdata$Status, levels = poststatus)

data_poststatus_tally <- 
  postdata %>% 
  group_by(Status) %>% 
  tally() %>% 
  filter(!is.na(Status)) 

ggplot(data_poststatus_tally, 
       aes(Status, y = 100 * (n/sum(n)),
           n)) +
  geom_bar(stat = "identity", fill="orange") +
  geom_text(aes(label=n), size= 4) +
  scale_x_discrete(labels = function(x) lapply(strwrap(x, width = 10, simplify = FALSE), paste, collapse="\n")) +
  theme_classic() +
  xlab("Status (Post-Survey)") +
  ylab("% Respondents") +
  ggtitle("Majority of Post-Survey Respondents were XXX") +
  theme(plot.title = element_text(hjust = 0.5)) +
  theme_classic(base_size = 14)

### KLJ Graduate Student is missing from the post-survey plot
```

```{r echo=FALSE}
### KLJ Breakdown of pre-survey respondents by discipline
 data_discipline_tally <- 
  predata %>% 
  group_by(Discipline) %>% 
  tally() %>% 
  mutate(`%` = round(n / sum(n) * 100, 1)) %>% 
  arrange(desc(n)) %>%
  filter(!is.na(Discipline))
colnames <- c("Discipline (Pre-Survey)", "n", "%")
kable(data_discipline_tally)

### KLJ Breakdown of post-survey respondents by research
 data_research_tally <- 
  postdata %>% 
  group_by(Research) %>% 
  tally() %>% 
  mutate(`%` = round(n / sum(n) * 100, 1)) %>% 
  arrange(desc(n)) %>%
  filter(!is.na(Research))
colnames <- c("Research (Post-Survey)", "n", "%")
kable(data_research_tally)
### KLJ How can we cut the research table off so that it doesn't show responses at 1% or less?
```

```{r echo=FALSE}
### KLJ Plot of programming usage pre-survey 
### KLJ How can we change the order of the factors? It needs to be:
### Not sure, I have never programmed, Less than once a year, Several times a year, Weekly, Monthly, Daily
data_programming_usage_tally <- 
  predata %>% 
  group_by(`Programming-Usage`) %>% 
  tally() %>% 
  filter(!is.na(`Programming-Usage`)) 

ggplot(data_programming_usage_tally, 
       aes(`Programming-Usage`, y = 100 * (n/sum(n)),
           n)) +
  geom_bar(stat = "identity", fill="orange") +
  geom_text(aes(label=n), size= 4) +
  scale_x_discrete(labels = function(x) lapply(strwrap(x, width = 10, simplify = FALSE), paste, collapse="\n")) +
  theme_classic() +
  xlab(" ") +
  ylab("% Respondents") +
  ggtitle("Programming Usage (Pre-Workshop)") +
  theme(plot.title = element_text(hjust = 0.5)) +
  theme_classic(base_size = 14)
```

```{r echo=FALSE}
### KLJ Plot of satisfaction with data management strategy and data analysis workflow (pre-survey)
### KLJ How can we put this information on one plot instead of two?
### KLJ How can we fix the order of the factors?
data_management_strategy_tally <- 
  predata %>% 
  group_by(`Data-Management-Strategy`) %>% 
  tally() %>% 
  filter(!is.na(`Data-Management-Strategy`)) 

ggplot(data_management_strategy_tally, 
       aes(`Data-Management-Strategy`, y = 100 * (n/sum(n)),
           n)) +
  geom_bar(stat = "identity", fill="orange") +
  geom_text(aes(label=n), size= 4) +
  scale_x_discrete(labels = function(x) lapply(strwrap(x, width = 10, simplify = FALSE), paste, collapse="\n")) +
  theme_classic() +
  xlab("") +
  ylab("% Respondents") +
  ggtitle("Perception of Pre-Workshop Data Management Strategy") +
  theme(plot.title = element_text(hjust = 0.5)) +
  theme_classic(base_size = 14)


data_analysis_workflow_tally <- 
  predata %>% 
  group_by(`Data-Analysis-Workflow`) %>% 
  tally() %>% 
  filter(!is.na(`Data-Analysis-Workflow`)) 

ggplot(data_analysis_workflow_tally, 
       aes(`Data-Analysis-Workflow`, y = 100 * (n/sum(n)),
           n)) +
  geom_bar(stat = "identity", fill="orange") +
  geom_text(aes(label=n), size= 4) +
  scale_x_discrete(labels = function(x) lapply(strwrap(x, width = 10, simplify = FALSE), paste, collapse="\n")) +
  theme_classic() +
  xlab(" ") +
  ylab("% Respondents") +
  ggtitle("Perception of Pre-Workshop Data Analysis Workflow") +
  theme(plot.title = element_text(hjust = 0.5)) +
  theme_classic(base_size = 14)
```
```{r echo=FALSE}
### KLJ Code chunk for Likert items (pre-survey)
cols_with_Agree <- map_lgl(predata, ~`%in%`("Agree", .x))
predata_agree <-  predata[ , cols_with_Agree]

levels = c("Strongly disagree", "Disagree", "Neutral", "Agree", "Strongly agree")

 factorfunction <- function(predata, factlevel){
  factor(predata, 
         levels=factlevel, 
         ordered = TRUE)
    fct_unify(predata, 
              levels=factlevel)}

predata_agree_likert <- likert(data.frame(lapply(predata_agree, factor, levels, ordered=TRUE)))
 
title <- "Pre-Workshop Perception of Tools"
 theme_update(plot.title = element_text(hjust = 0.5))
plot(predata_agree_likert) + ggtitle(title) 

 predata_agree <- map_if(predata_agree,
                     is.character,
                     as.factor)
 ### KLJ How can we change the axis lables?
```

```{r echo=FALSE}
 ### KLJ Do learners come to the workshop having a data set?
 data_dataset_tally <- 
  predata %>% 
  group_by(`Have-Dataset`) %>% 
  tally() %>% 
  mutate(`%` = round(n / sum(n) * 100, 1)) %>% 
  arrange(desc(n)) %>%
  filter(!is.na(`Have-Dataset`)) 
colnames <- c("Have Dataset?", "n", "%")
 kable(data_dataset_tally, col.names = colnames)
```

```{r echo=FALSE}
### KLJ Code chunk for Likert items (post-survey)
cols_with_Agree <- map_lgl(postdata, ~`%in%`("Agree", .x))
postdata_agree <-  postdata[ , cols_with_Agree]

levels = c("Strongly disagree", "Disagree", "Neutral", "Agree", "Strongly agree")

 factorfunction <- function(postdata, factlevel){
  factor(postdata, 
         levels=factlevel, 
         ordered = TRUE)
    fct_unify(postdata, 
              levels=factlevel)}

postdata_agree_likert <- likert(data.frame(lapply(postdata_agree, factor, levels, ordered=TRUE)))
 
title <- "Post-Workshop Perception of Atmosphere & Tools"
 theme_update(plot.title = element_text(hjust = 0.5))
plot(postdata_agree_likert) + ggtitle(title) 

 postdata_agree <- map_if(postdata_agree,
                     is.character,
                     as.factor)
 ### KLJ How can we change the axis lables?
 ### KLJ Can we make two separate Likert plots, one about workshop atmosphere and one for tools?
```

```{r echo=FALSE}
### KLJ Code chunk for likert plot (post-workshop perception of workshop instructors/helpers)
### KLJ This plot isn't working for some reason.
ordered_often <-
  c("Never", "Rarely", "Sometimes", "Often", "All of the time")

# KLJ Instructors
likert_cols_often_Inst <- 
 postdata %>% 
   select(grep("Instructors-", names(.))) %>% 
   mutate_if(is.character, as.factor) %>% 
   mutate_all(funs(fct_relevel(., ordered_often))) %>% 
   filter_all(all_vars(!is.na(.))) 
   
names(likert_cols_often_Inst) <- 
  gsub("Instructors-", "", names(likert_cols_often_Inst))
  
lc_often_inst <- likert(data.frame(likert_cols_often_Inst))

plot(lc_often_inst) +
  ggtitle("Perception of Workshop Instructors")

# KLJ Double check the plot with a table:
xx <- 
likert_cols_often_Inst %>% 
  gather(key, value) %>% 
  group_by(key, value) %>% 
  tally() %>% 
  mutate(perc = round(n / sum(n) * 100, 1)) %>% 
  select(-n) %>% 
  spread(key, perc) %>% 
  slice(match(ordered_often,
              value))

# KLJ Helpers
likert_cols_often_h <- 
 postdata %>% 
   select(grep("Helpers-", names(.))) %>% 
   mutate_if(is.character, as.factor) %>% 
   mutate_all(funs(fct_relevel(., ordered_often))) %>% 
   filter_all(all_vars(!is.na(.))) 
   
names(likert_cols_often_h) <- 
  gsub("Helpers-", "", names(likert_cols_often_h))
  
lc_often_h <- likert(data.frame(likert_cols_often_h))

plot(lc_often_h) +
  ggtitle("Perception of Workshop Helpers")

# KLJ Double check the plot with a table:
xx <- 
likert_cols_often_h %>% 
  gather(key, value) %>% 
  group_by(key, value) %>% 
  tally() %>% 
  mutate(perc = round(n / sum(n) * 100, 1)) %>% 
  select(-n) %>% 
  spread(key, perc) %>% 
  slice(match(ordered_often,
              value))
```